\pkg{ggenealogy}: An \proglang{R} Package for Visualizing Genealogical Data

### 2016-07-14 article.tex
abstract

%% an abstract and keywords
\Abstract{
  This paper introduces \pkg{ggenealogy} (\citealt{ggen}), a developing \proglang{R} software package that provides tools for searching through genealogical data, generating basic statistics on their graphical structures using parent and child connections, and displaying the results. It is possible to draw the genealogy in relation to variables related to the nodes, and to determine and display the shortest path distances between the nodes. Production of pairwise distance matrices and genealogical diagrams constrained on generation are also available in the visualization toolkit. The tools are being tested on a dataset with milestone cultivars of soybean varieties (\citealt{soybean}) as well as on a web-based database of the academic genealogy of mathematicians (\citealt{mgp}). The software package is currently available on the Comprehensive \proglang{R} Archive Network.}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> install.packages("ggenealogy")
R> library("ggenealogy")
R> data("sbGeneal")
R> str(sbGeneal)
\end{CodeInput}
\begin{CodeOutput}
'data.frame':	390 obs. of  5 variables:
 $ child       : chr  "5601T" "Adams" "A.K." "A.K. (Harrow)" ...
 $ year        : num  1981 1948 1910 1912 1968 ...
 $ yield       : int  NA 2734 NA 2665 NA 2981 2887 2817 NA NA ...
 $ year.imputed: logi  TRUE FALSE TRUE FALSE FALSE FALSE ...
 $ parent      : chr  "Hutcheson" "Dunfield" NA "A.K." ...
\end{CodeOutput}
\end{CodeChunk}

\subsection{Academic genealogy of statisticians}

The \pkg{ggenealogy} package also comes with an academic genealogy of statisticians; this dataset is in the form of a data frame with 8165 rows and six columns. To develop this later dataset, we contacted the \citealt{mgp}, a web-based database for the genealogy of academic mathematicians. This database, which currently contains almost 200,000 entries, is a service of the North Dakota State University Department of Mathematics and the American Mathematical Society. The Mathematics Genealogy Project contact provided us a Structured Query Language (SQL) export, and we used PostgreSQL to query the database (\citealt{psql}).

@@

We now convert it into an \pkg{igraph} object (\citealt{igraph}) \code{sbIG} using the function \code{dfToIG()}.

\begin{CodeChunk}
\begin{CodeInput}
R> sbIG <- dfToIG(sbGeneal)
R> sbIG
\end{CodeInput}
\begin{CodeOutput}
IGRAPH UNW- 230 340 -- 
+ attr: name (v/c), weight (e/n)
+ edges (vertex names):
 [1] 5601T    --Hutcheson        Adams    --Dunfield        
 [3] A.K.     --A.K. (Harrow)    Altona   --Flambeau        
 [5] Amcor    --Amsoy 71         Adams    --Amsoy           
 [7] Amsoy 71 --C1253            Anderson --Lincoln         
 [9] Bay      --York             Bedford  --Forrest         
[11] Beeson   --Kent             Blackhawk--Richland        
[13] Bonus    --C1266R           Bradley  --J74-39          
[15] Bragg    --Jackson          Bragg    --Bragg x D60-7965
+ ... omitted several edges
\end{CodeOutput}
\end{CodeChunk}

@@

$numComponents
[1] 11

$avePathLength
[1] 5.333746

$graphDiameter
[1] 13

$numNodes
[1] 230

$numEdges
[1] 340

$logN
[1] 5.438079
\end{CodeOutput}
\end{CodeChunk}

>>>>

%% an abstract and keywords
\Abstract{
  This paper introduces \pkg{ggenealogy} (\citealt{ggen}), a developing \proglang{R} software package that provides tools for searching through genealogical data, generating basic statistics on their graphical structures using parent and child connections, and displaying the results. It is possible to draw the genealogy in relation to variables related to the nodes, and to determine and display the shortest path distances between the nodes. Production of pairwise distance matrices and genealogical diagrams constrained on generation are also available in the visualization toolkit. The tools are being tested on a dataset with milestone cultivars of soybean varieties (\citealt{soybean}) as well as on a web-based database of the academic genealogy of mathematicians (\citealt{mgp}). The latest stable package version is available in source and binary form on the Comprehensive \proglang{R} Archive Network.}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> install.packages("ggenealogy")
R> library("ggenealogy")
R> data("sbGeneal")
R> str(sbGeneal)
\end{CodeInput}
\begin{CodeOutput}
'data.frame':	390 obs. of  5 variables:
 $ child       : chr  "5601T" "Adams" "A.K." "A.K. (Harrow)" ...
 $ year        : num  1981 1948 1910 1912 1968 ...
 $ yield       : int  NA 2734 NA 2665 NA 2981 2887 2817 NA NA ...
 $ year.imputed: logi  TRUE FALSE TRUE FALSE FALSE FALSE ...
 $ parent      : chr  "Hutcheson" "Dunfield" NA "A.K." ...
\end{CodeOutput}
\end{CodeChunk}
\subsection{Academic genealogy of statisticians}
The \pkg{ggenealogy} package also comes with an academic genealogy of statisticians; this dataset is in the form of a data frame with 8165 rows and six columns. To develop this later dataset, we contacted the \citealt{mgp}, a web-based database for the genealogy of academic mathematicians. This database, which currently contains almost 200,000 entries, is a service of the North Dakota State University Department of Mathematics and the American Mathematical Society. The Mathematics Genealogy Project contact provided us a Structured Query Language (SQL) export, and we used PostgreSQL to query the database (\citealt{psql}).

@@

We now convert it into an \pkg{igraph} object (\citealt{igraph}) \code{sbIG} using the function 
\code{dfToIG()}.
\begin{CodeChunk}
\begin{CodeInput}
R> sbIG <- dfToIG(sbGeneal)
R> sbIG
\end{CodeInput}
\begin{CodeOutput}
IGRAPH UNW- 230 340 -- 
+ attr: name (v/c), weight (e/n)
+ edges (vertex names):
 [1] 5601T    --Hutcheson        Adams    --Dunfield        
 [3] A.K.     --A.K. (Harrow)    Altona   --Flambeau        
 [5] Amcor    --Amsoy 71         Adams    --Amsoy           
 [7] Amsoy 71 --C1253            Anderson --Lincoln         
 [9] Bay      --York             Bedford  --Forrest         
[11] Beeson   --Kent             Blackhawk--Richland        
[13] Bonus    --C1266R           Bradley  --J74-39          
[15] Bragg    --Jackson          Bragg    --Bragg x D60-7965
+ ... omitted several edges
\end{CodeOutput}
\end{CodeChunk}

@@

$numComponents
[1] 11
$avePathLength
[1] 5.333746

$graphDiameter
[1] 13
$numNodes
[1] 230

$numEdges
[1] 340
$logN
[1] 5.438079
\end{CodeOutput}
\end{CodeChunk}
